---
title: "ExtraneousWork"
author: "jsalzwedel"
date: "05/27/2015"
output: html_document
---


We are interested in the health and economic impacts of events. Let's subset the
relevant sections of the data: fatalities, injuries, property damage, and crop 
damage, along with the event type and the event reference number.  Rather than 
judge the relative severity of X fatalities vs Y injuries, let's combine both of 
those into one category called Casualties.  Likewise with the property and
crop damage.

```{r}
impact <- data.table(EventType = data$EVTYPE,
                     Casualties = data$FATALITIES + data$INJURIES,
                     Costs = data$PROPDMG + data$CROPDMG,
                     ReferenceNum = data$REFNUM)
```




Let's find the event types where the most casualties occured.  Rather than look
at individual events, let's look at the average casualties for a given event
type.

```{r}
avgCasualties <- aggregate(x = list(AvgCasualties = impact$Casualties),
                           by = list(EventType = impact$EventType),
                           FUN = mean)
avgCasualties <- data.table(avgCasualties)
# Sort to list the event types with most casualties first, then remove event
# types that have no casualties
setorder(avgCasualties,-AvgCasualties)
avgCasualties <- avgCasualties[ which(avgCasualties$AvgCasualties > 0), ]
```



```{r}
# Clean up naming variations
```


```{r}
impact$EventType <- tolower(impact$EventType)
```



# Exploratory work
```{r}
with(data,plot(EVTYPE,FATALITIES))
avgFatalities <- aggregate(x = list(AvgFatalities = data$FATALITIES),
                           by=list(Event = data$EVTYPE),
                           FUN=mean)
avgFatalities <- data.table(avgFatalities)
setorder(avgFatalities,-AvgFatalities)
head(avgFatalities)
```

Let's take a closer look at the events where fatalities occured
```{r}
fatalData <- data[data$FATALITIES > 0,]
setorder(fatalData,-FATALITIES)
```


```{r}
avgInjuries <- aggregate(x = list(AvgInjuries = data$INJURIES),
                           by=list(Event = data$EVTYPE),
                           FUN=mean)
avgInjuries <- data.table(avgInjuries)
setorder(avgInjuries, -AvgInjuries)
head(avgInjuries,20)
```
