---
title: 'Reproducible Research: Peer Assessment 2'
author: "jsalzwedel"
date: "05/20/2015"
output: 
    html_document:
        keep_md: true
---

# Synopsis
This analysis will explore the NOAA Storm Database to look at the health impacts
and economic consequences of of severe weather events.


# Data Processing

## Acquiring and Reading the Data
```{r, cache=TRUE}
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(url,"StormData.csv.bz2",method="wget")
system("bunzip2 StormData.csv.bz2")
```

Load relevant libraries
```{r}
library(data.table)
```

Read in the data
```{r, cache=TRUE}
data <- read.csv("StormData.csv")
data <- data.table(data)
```


## Cleaning

### Tidying Event Type

The EVTYPE category is a mess.  Many event types don't follow the naming
conventions set out in the codebook.  For example, there are a lot of events
that are effectively the same event type, but have slightly different names (e.g
"thunderstorm wind", "tstm wind", and "thunderstormw").  We should simplify all 
these to the same name.  Furthermore, there are also a lot of events with the
same event type but different capitalization schemes (e.g. "HEAT WAVE" and "Heat
Wave").  We can catch and clean up these classification mismatches by making all
the event types lower case.  In the end, all event types should fall into one of
the following approved categories:

```{r}
eventTypes <- c("Astronomical Low Tide", "Avalanche", "Blizzard", 
                "Coastal Flood", "Cold/Wind Chill", "Debris Flow", "Dense Fog",
                "Dense Smoke", "Drought", "Dust Devil", "Dust Storm", 
                "Excessive Heat", "Extreme Cold/Wind Chill", "Flash Flood", 
                "Flood", "Frost/Freeze", "Funnel Cloud", "Freezing Fog", "Hail", 
                "Heat", "Heavy Rain", "Heavy Snow", "High Surf", "High Wind", 
                "Hurricane (Typhoon)", "Ice Storm", "Lake-Effect Snow", 
                "Lakeshore Flood", "Lightning", "Marine Hail", 
                "Marine High Wind", "Marine Strong Wind", 
                "Marine Thunderstorm Wind", "Rip Current", "Seiche", "Sleet", 
                "Storm Surge/Tide", "Strong Wind", "Thunderstorm Wind", 
                "Tornado", "Tropical Depression", "Tropical Storm", "Tsunami",
                "Volcanic Ash", "Waterspout", "Wildfire", "Winter Storm", 
                "Winter Weather")
```

To start, let's make all the event types lower case.

```{r}
data$EVTYPE <- tolower(data$EVTYPE)
eventTypes <- tolower(eventTypes)
```


### Tidying property/crop damage

Now let's update the property and crop damage numbers to all be in dollars,
rather than thousands/millions/billions of daughters, as described by the 
*DMGEXP columns.









# Analysis




# Results








