---
title: 'Reproducible Research: Peer Assessment 2'
author: "jsalzwedel"
date: "05/20/2015"
output: 
    html_document:
        keep_md: true
---

# Synopsis
This analysis will explore the NOAA Storm Database to look at the health impacts
and economic consequences of of severe weather events.


# Data Processing
```{r, cache=TRUE}
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
download.file(url,"StormData.csv.bz2",method="wget")
system("bunzip2 StormData.csv.bz2")
```

Load relevant libraries
```{r}
library(data.table)
library(tis)
```

Read in the data
```{r, cache=TRUE}
data <- read.csv("StormData.csv")
data <- data.table(data)
```

We are interested in the health and economic impacts of events. Let's subset the
relevant sections of the data: fatalities, injuries, property damage, and crop 
damage, along with the event type and the event reference number.

```{r}
subsetData <- subset(data,select=c(EVTYPE,FATALITIES,INJURIES,PROPDMG,CROPDMG,REFNUM))
```

Rather than judge the relative severity of X fatalities vs Y injuries, let's
combine both of those into one category called Casualties

```{r}
impact <- data.table(EventType = data$EVTYPE,
                     Casualties = data$FATALITIES + data$INJURIES,
                     Costs = data$PROPDMG + data$CROPDMG,
                     ReferenceNum = data$REFNUM)
```

```{r}
with(data,plot(EVTYPE,FATALITIES))
avgFatalities <- aggregate(x = list(AvgFatalities = data$FATALITIES),
                           by=list(Event = data$EVTYPE),
                           FUN=mean)
avgFatalities <- data.table(avgFatalities)
setorder(avgFatalities,-AvgFatalities)
head(avgFatalities)
```

Let's take a closer look at the events where fatalities occured
```{r}
fatalData <- data[data$FATALITIES > 0,]
setorder(fatalData,-FATALITIES)
```


```{r}
avgInjuries <- aggregate(x = list(AvgInjuries = data$INJURIES),
                           by=list(Event = data$EVTYPE),
                           FUN=mean)
avgInjuries <- data.table(avgInjuries)
setorder(avgInjuries, -AvgInjuries)
head(avgInjuries,20)
```

# Results